---
title: "hw1"
author: "Zihao_Wang"
date: "10/7/2018"
output: html_document
---

```{r}
rm(list=ls())
set.seed(12345)
```


# 1. Computing and plotting with R

## (a)
```{r}
## function to compute empirical mean square error 
mse <- function(n,sigma){
  mysample = rnorm(n = n, sd = sigma, mean = 1)
  return((mean(mysample)-1)^2)
}

emse <- function(n,sigma,B){
  return(mean(replicate(B,mse(n,sigma))))
}

## simulate and plot
sigma = 1
B = 100

ns = seq(1,1000,5)
results = replicate(length(ns),0)
theory = replicate(length(ns),0)

for(i in 1:length(ns)){
  results[i] = emse(ns[i],sigma,B)
  theory[i] = sigma^2/ns[i]
}

plot(log(ns), log(results), xlab = "n", ylab = "mse")
lines(log(ns),log(theory), col = "red")
```

### Comment:
From the simulation experiments, we can see the results align with theoretical results

## (b)

```{r}
n = 1000
sigma = 1
B = 100

## simulation
Z = replicate(B,sqrt(n)*(mean(rnorm(n,1,sigma)) - 1))

## theoretical standard normal
ns = seq(-10,10,0.01)
theory = replicate(length(ns),0)
for(i in 1:length(ns)){
  theory[i] = 1/sqrt(2*pi) * exp(-ns[i]^2 *0.5)
}

plot(density(Z,), col = "blue")
lines(ns,theory,col = "red")
```


# 2 Leave-oue-out cross-validation

## (a)
By definition, $\hat{R}(h) = \frac{1}{n} * \sum_{i = 1}^{n} (r(x_{i})-\hat{r}_{-i}(x_{i}))^2$. 
Then we have $r(x_{i}) - \hat{r}_{-i}(x_{i}) = Y_{i} - \frac{\sum_{k \neq i} L_{i,k} * Y_{k}}{1-L_{ii}} = \frac{Y_{i} - \hat{r}_{n}(x_{i})} {1-L  _{ii}}$. Then our desired equation follows.

## (b)
```{r}
StochasticDoppler <- function(x,sigma=0.1){
  y = sqrt(x*(1-x)) * sin(2.1*pi/(x+0.05)) + sigma * rnorm(1,0,1)
  return(y)
}

Doppler <- function(x){
  y = sqrt(x*(1-x)) * sin(2.1*pi/(x+0.05)) 
  return(y)
}

N = 1000
Xs = seq(1,N)/N
ys <- sapply(Xs, StochasticDoppler)
ys_true = sapply(Xs,Doppler)
plot(Xs,ys)

```


```{r}
## Gaussian kernel function
K <- function(x){
  return(1/sqrt(2*pi) * exp(-x^2/2))
}

# L_ii = K(0)/SUM(K((Xi-Xj)/h))
L_ii <- function(x,Xs,h){
  Lii = K(0)/sum(sapply(Xs, function(Xj) K((Xj-x)/h)))
  return(Lii)
}


degree_freedom <- function(Xs,h){
  ys_hat = ksmooth(Xs,ys,kernel = "normal", bandwidth = h)$y
  Liis = sapply(Xs,function(Xi) L_ii(Xi,Xs,h))
  return(sum(Liis))
}

LOOCV <- function(Xs,ys,ys_hat,h){
  n = length(Xs)
  NUMERATOR = ys - ys_hat
  DENOMINATOR = 1- sapply(Xs,function(Xi) L_ii(Xi,Xs,h))
  R_hat = 1/n * sum((NUMERATOR/DENOMINATOR)^2)
  return(R_hat)
}

SmoothLOOCV <- function(Xs,ys,h){
  ys_hat = ksmooth(Xs,ys,kernel = "normal", bandwidth = h)$y
  return(LOOCV(Xs,ys,ys_hat,h))
}


# hs = seq(1,10,1)/1000
# loocv = sapply(hs, function(h) SmoothLOOCV(Xs,ys,h))
# plot(hs,loocv)

# dfs = sapply(hs,function(h) degree_freedom(Xs,h))
# plot(loocv,dfs)





```

```{r}
y_hat = ksmooth(Xs,ys,kernel = "normal", bandwidth  = 0.005)$y
plot(Xs,ys_true, type = "l", col = "red")
lines(Xs,y_hat,type = "l", col = "blue")

```


# 3. Kernel density estimate for Old Faithful Geyser
```{r}
data("faithful")
plot(density(faithful$eruptions))
plot(density(faithful$waiting))
```


$\hat{J}_{n}(h) = \int \hat{f}_{n}(x) dx - 2/n * \sum_{i = 1}^{n} \hat{f}_{n}^{(-i)} (x_{i})$, if we call the second half 2* mean_pdf,
 then mean_pdf can be written as: $\frac{1}{n(n-1)h} \sum_{i} \sum_{j \neq i} K(\frac{x_i - x_j}{h} )$
```{r}
## f_hat uses Gaussian kernel to estimate pdf, given
### Xs: a vector of all observations and h: the bandwidth
f_hat <- function(x, Xs,h){
  n = length(Xs)
  scaled_Xs = sapply(Xs, function(xi) (xi-x)/h)
  p = 1/n* 1/h * sum(sapply(scaled_Xs,K))
  return(p)
}


## LOOCV computes loocv using the formula above
LOOCV <- function(Xs,h){
  ## compute integral
  f_hat_sqr = function(x) f_hat(x,Xs,h)^2
  integral = integrate(Vectorize(f_hat_sqr),0,10,subdivisions = 500)$value
  ## comnpute the mean pdf for 
  n = length(Xs)
  X_matrix = t(replicate(n,Xs))
  ## for ith row of X_matrix, subtract X_i
  X_matrix_scaled  = (X_matrix - t(X_matrix)) / h
  X_matrix_kernel = structure(sapply(X_matrix_scaled,K), dim = dim(X_matrix_scaled))
  mean_pdf = 1/(n*(n-1)*h) * sum(X_matrix_kernel) - sum(diag(X_matrix_kernel))
  
  return(integral - 2*mean_pdf)
}


# ## for eruptions
# hs0 = seq(1,10,1)/10000
# loocvs0 = sapply(hs0, function(h) LOOCV(faithful$eruptions,h))
# 
# hs = seq(1,10,1)/1000
# loocvs = sapply(hs, function(h) LOOCV(faithful$eruptions,h))
# 
# hs2 = seq(1,10,1)/100
# loocvs2 = sapply(hs2, function(h) LOOCV(faithful$eruptions,h))
# 
# hs3 = seq(1,10,1)/10
# loocvs3 = sapply(hs3, function(h) LOOCV(faithful$eruptions,h))
# 
# hss = c(hs0,hs,hs2,hs3)
# loocvss = c(loocvs0,loocvs,loocvs2,loocvs3)
# plot(hss, loocvss)
# # xs = seq(0,10,0.1)
# # plot(xs,sapply(xs, function(x) f_hat(x,faithful$eruptions,0.05)), type = "l")
# 
# ## for eruptions loocv seems to be problematic: increasing with h; drop to negative when h < 1e-05
# 
# ## for waiting time
# loocvs_waiting0 = sapply(hs0, function(h) LOOCV(faithful$waiting,h))
# plot(hs0,loocvs_waiting0)
# 
# loocvs_waiting = sapply(hs, function(h) LOOCV(faithful$waiting,h))
# 
# loocvs_waiting2 = sapply(hs2, function(h) LOOCV(faithful$waiting,h))
# 
# loocvs_waiting3 = sapply(hs3, function(h) LOOCV(faithful$waiting,h))
# plot(hss, c(loocvs_waiting,loocvs_waiting0,loocvs_waiting2,loocvs_waiting3))

```

##  plot estimated density with optimum h
```{r}
h_star_eruption = 0.3
xs = seq(0,10,0.1)
plot(xs,sapply(xs, function(x) f_hat(x,faithful$eruptions,h_star_eruption)), type = "l")

h_star_waiting = 1.5
plot(xs,sapply(xs, function(x) f_hat(x,faithful$waiting,h_star_waiting)), type = "l")
```


```{r}
library(kedd)
ucv_eruptions = h.ucv(faithful$eruptions,nb = 1000)
plot(ucv_eruptions)

plot(xs,sapply(xs, function(x) f_hat(x,faithful$eruptions,ucv_eruptions$h)), type = "l")
```

```{r}
ucv_waiting = h.ucv(faithful$waiting,nb = 1000)
plot(ucv_waiting)

plot(xs,sapply(xs, function(x) f_hat(x,faithful$waiting,ucv_eruptions$h)), type = "l")
```





